Case Study (McKenna)

### 2020-03-10 Code/mrp/Case Study (McKenna) .Rmd
Splitting up the case study
Cleaning up the code.

---
title: "MRP Case Study"
author: "McKenna Weech, Adam Hettinger, and Hannah Thompson"
date: "1/17/2020"
output: github_document
---

@@

# Hierarchal Logit Model

```{}

 data {
   // Define variables in data
   // Number of observations (an integer)
   int<lower=0> N;
   // Number of parameters
   int<lower=0> p;
   // Variables
   int death[N];
   int<lower=0>  qsmk[N];
   int<lower=0>  sex[N];
   real<lower=0> age[N];
   int<lower=0>  race[N];
   real<lower=0> smokeyrs[N];
 }

 parameters {
   // Define parameters to estimate
   real beta[p];
 }

 transformed parameters  {
   // Probability trasformation from linear predictor
   real<lower=0> odds[N];
   real<lower=0, upper=1> prob[N];

   for (i in 1:N) {
     odds[i] <- exp(beta[1] + beta[2]*qsmk[i] + beta[3]*sex[i] + beta[4]*age[i] + beta[5]*race[i] + beta[6]*smokeyrs[i]);
     prob[i] <- odds[i] / (odds[i] + 1);
   }
 }

 model {
   // Prior part of Bayesian inference (flat if unspecified)

  // Likelihood part of Bayesian inference
    death ~ bernoulli(prob);
}

```{r echo=FALSE}
# 
# J <- c(2, 3, 7, 3, 50) # male or not, eth, age, income level, state
#   poststrat <- as.data.frame(array(NA, c(prod(J), length(J)+1))) # Columns of post-strat matrix, plus one for size
#   colnames(poststrat) <- c("male", "eth", "age","income", "state",'N')
#   count <- 0
#   for (i1 in 1:J[1]){
#     for (i2 in 1:J[2]){
#       for (i3 in 1:J[3]){
#         for (i4 in 1:J[4]){
#           for (i5 in 1:J[5]){
#               count <- count + 1
#               # Fill them in so we know what category we are referring to
#               poststrat[count, 1:5] <- c(i1-1, i2, i3, i4, i5)
#           }
#         }
#       }
#     }
#   }
#   # Proportion in each sample in the population
#   p_male <- c(0.52, 0.48)
#   p_eth <- c(0.5, 0.2, 0.3)
#   p_age <- c(0.2,.1,0.2,0.2, 0.10, 0.1, 0.1)
#   p_income<-c(.50,.35,.15)
#   p_state_tmp<-runif(50,10,20)
#   p_state<-p_state_tmp/sum(p_state_tmp)
#   poststrat$N<-0
#   for (j in 1:prod(J)){
#     poststrat$N[j] <- round(250e6 * p_male[poststrat[j,1]+1] * p_eth[poststrat[j,2]] *
#       p_age[poststrat[j,3]]*p_income[poststrat[j,4]]*p_state[poststrat[j,5]]) #Adjust the N to be the number observed in each category in each group
#   }
# 
#   # Now let's adjust for the probability of response
#   p_response_baseline <- 0.01
#   p_response_male <- c(2, 0.8) / 2.8
#   p_response_eth <- c(1, 1.2, 2.5) / 4.7
#   p_response_age <- c(1, 0.4, 1, 1.5,  3, 5, 7) / 18.9
#   p_response_inc <- c(1, 0.9, 0.8) / 2.7
#   p_response_state <- rbeta(50, 1, 1)
#   p_response_state <- p_response_state / sum(p_response_state)
#   p_response <- rep(NA, prod(J))
#   for (j in 1:prod(J)) {
#     p_response[j] <-
#       p_response_baseline * p_response_male[poststrat[j, 1] + 1] *
#       p_response_eth[poststrat[j, 2]] * p_response_age[poststrat[j, 3]] *
#       p_response_inc[poststrat[j, 4]] * p_response_state[poststrat[j, 5]]
#   }
# 
#   coef_male <- c(0,-0.3)
#   coef_eth <- c(0, 0.6, 0.9)
#   coef_age <- c(0,-0.2,-0.3, 0.4, 0.5, 0.7, 0.8, 0.9)
#   coef_income <- c(0,-0.2, 0.6)
#   coef_state <- c(0, round(rnorm(49, 0, 1), 1))
#   coef_age_male <- t(cbind(c(0, .1, .23, .3, .43, .5, .6),
#                            c(0, -.1, -.23, -.5, -.43, -.5, -.6)))
#   true_popn <- data.frame(poststrat[, 1:5], service_failure = rep(NA, prod(J)))
#   for (j in 1:prod(J)) {
#     true_popn$satisfaction[j] <- plogis(
#       coef_male[poststrat[j, 1] + 1] +
#         coef_eth[poststrat[j, 2]] + coef_age[poststrat[j, 3]] +
#         coef_income[poststrat[j, 4]] + coef_state[poststrat[j, 5]] +
#         coef_age_male[poststrat[j, 1] + 1, poststrat[j, 3]]
#       )
#   }
```

Now that we have our simulated data, we now need our hierarchal stan model.

```{}

// Index values, observations, and covariates.
data {
  int<lower = 1> N;                      // Number of observations.
  int<lower = 1> K;                      // Number of groups.
  int<lower = 1> I;                      // Number of observation-level covariates.

  real service_failure[N];                        // Vector of observations.
  int<lower = 1, upper = K> male[N];        // Vector of group assignments.
  int age[N];                               // Vector of ethnicity covariates.
  int eth[N];                               // Vector of age covariates.
  int income[N];                          // Vector of income covariates.
  int state[N];                          // Vector of state covariates.

  real gamma_mean;                       // Mean for the hyperprior on gamma.
  real<lower = 0> gamma_var;             // Variance for the hyperprior on gamma.
  real<lower = 0> tau_min;               // Minimum for the hyperprior on tau.
  real<lower = 0> tau_max;               // Maximum for the hyperprior on tau.
  real<lower = 0> sigma_min;             // Minimum for the hyperprior on tau.
  real<lower = 0> sigma_max;             // Maximum for the hyperprior on tau.
}

// Parameters and hyperparameters.
parameters {
  matrix[K, (I - 1)] alpha;              // Matrix of observation-level gender coefficients.
  vector[K] beta;                        // Vector of all other observation-level coefficients.
  real gamma;                            // Mean of the population model.
  real<lower=0> tau;                     // Variance of the population model.
  real<lower=0> sigma;                   // Variance of the observation model.
}

// Hierarchical regression.
model {
  // Declare mu for use in the linear model.
  vector[N] mu;

  // Hyperpriors and prior.
  gamma ~ normal(gamma_mean, gamma_var);
  tau ~ uniform(tau_min, tau_max);
  sigma ~ uniform(sigma_min, sigma_max);

  // Population model and likelihood.
  for (k in 1:K) {
    alpha[k,] ~ normal(gamma, tau);
    beta[k] ~ normal(gamma, tau);
  }
  for (n in 1:N) {
    mu[n] = alpha[male[n], age[n]] + beta[male[n]] * eth[n] + beta[male[n]] * 
    income[n] + beta[male[n]] * state[n];
  }
  service_failure ~ normal(mu, sigma);
}

// Generate predictions using the posterior.
generated quantities {
  vector[N] mu_pc;                       // Declare mu for predicted linear model.
  real service_failure_pc[N];                     // Vector of predicted observations.

  // Generate posterior prediction distribution.
  for (n in 1:N) {
    mu_pc[n] = alpha[male[n], age[n]] + beta[male[n]] * eth[n] + beta[male[n]] * 
    income[n] + beta[male[n]] * state[n];
    service_failure_pc[n] = normal_rng(mu_pc[n], sigma);
  }
}

As we will also need to do some prior and posterior checks, and for that we will need the generative model.

```{}

// Index values, observations, covariates, and hyperior values.
data {
  int<lower = 1> N;                      // Number of observations.
  int<lower = 1> K;                      // Number of groups.
  int<lower = 1> I;                      // Number of observation-level covariates.

  int<lower = 1, upper = K> male[N];        // Vector of group assignments.
  int eth[N];                               // Vector of ethnicity covariates.
  int age[N];                               // Vector of age covariates.
  int income[N];                          // Vector of income covariates.
  int state[N];                          // Vector of state covariates.

  real gamma_mean;                       // Mean for the hyperprior on gamma.
  real<lower = 0> gamma_var;             // Variance for the hyperprior on gamma.
  real<lower = 0> tau_min;               // Minimum for the hyperprior on tau.
  real<lower = 0> tau_max;               // Maximum for the hyperprior on tau.
  real<lower = 0> sigma_min;             // Minimum for the hyperprior on tau.
  real<lower = 0> sigma_max;             // Maximum for the hyperprior on tau.
}

// Generate data according to the hierarchical regression.
generated quantities {
  matrix[K, (I - 1)] alpha;              // Matrix of observation-level gender coefficients.
  vector[K] beta;                        // Vector of all other observation-level coefficients.
  real gamma;                            // Mean of the population model.
  real<lower=0> tau;                     // Variance of the population model.
  real<lower=0> sigma;                   // Variance of the observation model.

  vector[N] mu;                          // Declare mu for linear model.
  real service_failure[N];                        // Vector of observations.

  gamma = normal_rng(gamma_mean, gamma_var);
  tau = uniform_rng(tau_min, tau_max);
  sigma = uniform_rng(sigma_min, sigma_max);

  // Draw parameter values and generate data.
  for (k in 1:K) {
    for (i in 1:(I - 1)) {
      alpha[k, i] = normal_rng(gamma, tau);
    }
    beta[k] = normal_rng(gamma, tau);
  }
  for (n in 1:N) {
    mu[n] = alpha[male[n], eth[n]] + beta[male[n]] * age[n] + beta[male[n]] * 
    income[n] + beta[male[n]] * state[n];
    service_failure[n] = normal_rng(mu[n], sigma);
  }
}

@@

```{r echo=FALSE}

# # Specify data.
# data <- list(
#   N = nrow(true_popn),                     # Number of observations.
#   K = 2,                    # Number of groups.
#   I = 65,                                # Number of observation-level covariates.
#   
#   cat_pref = true_popn$cat_pref,               # Vector of observations.
#   male = as.numeric(as.factor(true_popn$male)),            # Vector of group assignments.
#   eth = true_popn$eth,       # Vector of brands covariates.
#   age = true_popn$age,        # Vector of price covariates.
#   income = true_popn$income,
#   state = true_popn$state,
#   
#   gamma_mean = 0.5,                                # Mean for the hyperprior on gamma.
#   gamma_var = .2,                                 # Variance for the hyperprior on gamma.
#   tau_min = 0,                                   # Minimum for the hyperprior on tau.
#   tau_max = 1,                                   # Maximum for the hyperprior on tau.
#   sigma_min = 0,                                 # Minimum for the hyperprior on tau.
#   sigma_max = 1                                 # Maximum for the hyperprior on tau.
# )
# # Calibrate the model.
# model03 <- stan(
#   file = here::here("Code/mrp/hierarchical_model.stan"),
#   data = data,
#   control = list(adapt_delta = 0.99),
#   seed = 42
# )

```{r echo=FALSE}
# Posterior predictive check.
# post_pc03 <- tibble(
#   # Extract the posterior predicted values.
#   intent = as.vector(extract(model03)$cat_pref_pc)
# )
# # Plot the posterior predictive distribution.
# ggplot(post_pc03, aes(x = intent)) +
#   geom_histogram() +
#   xlim(1, 10) 
# ggplot(purchase_intent, aes(x = intent)) +
#   geom_histogram() +
#   xlim(1, 10) 
```

```{r echo=FALSE}
# model03 %>%
#   gather_draws(alpha[n, i]) %>%
#   unite(.variable, .variable, n, i) %>%
#   ggplot(aes(x = .value, y = .variable)) +
#   geom_halfeyeh(.width = .95) +
#   facet_wrap(
#     ~ .variable,
#     nrow = data$K,
#     ncol = (data$I - 1),
#     scales = "free"
#   )
```

```{r echo=FALSE}
# Plot the betas.
# model03 %>%
#   gather_draws(beta[i]) %>%
#   unite(.variable, .variable, i) %>%
#   ggplot(aes(x = .value, y = .variable)) +
#   geom_halfeyeh(.width = .95) +
#   facet_wrap(
#     ~ .variable,
#     nrow = data$K,
#     ncol = 1,
#     scales = "free"
#   )
```

>>>>

---
title: "Case Study (McKenna)"
author: "McKenna Weech"
date: "3/10/2020"
output: html_document
---

@@

@@
